\begin{acknowledgments}
I would like to thank my advisor, Prof.~Dick Furnstahl, for his patient and steady guidance over the years. After giving me a starting point, he was consistently available to clarify concepts and give feedback on my work. As I worked my way up from not knowing anything about quantum mechanics, he helped guide my intuition and encouraged me to apply the technical skills I had acquired in my research. His knowledge, wisdom, productivity, and integrity inspire me daily to become a better scientist and a better member of society.

I am forever indebted to my parents, who have made everything I have accomplished possible. My father, Ulrich Heinz, helped to develop how I think about the world and solving problems in it. My mother, Christiane Heinz-Neidhart, has helped me develop discipline, work ethic, and a reasonable structure for my daily life. Their overwhelming support for my education and my ambitions is the reason I am where I am today.

I would also like to thank my brother, Michael Heinz, for inspiring and supporting me. His work ethic and brilliance consistently reminds me that there is more that I could do and better ways I could do it. I can count on him for a listening ear when I need it and honest, useful feedback and advice when I want it.

I want to thank the other members of my committee, Prof.~P Sadayappan and Prof.~Robert Perry, for their helpful input on my thesis. I want to thank the members of the low-energy nuclear theory group at OSU during the time I was here; their comments during group meetings, conferences, and isolated conversations have helped develop my knowledge of the field as a whole, which has also helped my own research.

Finally, I would like to thank all of my friends for their support. They have been understanding of my sometimes erratic availability and encouraging of my ambitions.
\end{acknowledgments}
